Keep Your Laziness In Check

### 2017-12-03 presentations/slides.tex
modified slides for 700 presentation

@@

@@

\begin{frame}[fragile]
\frametitle{Lazy lists}
\begin{minted}{ocaml}
type 'a lazyList =
  | Cons of ('a Lazy.t *
             'a lazyList Lazy.t)
  | Nil
\end{minted}
\end{frame}

@@

\note[itemize]{
  \item As a micro example to motivate why thinking about laziness can be
        non-trivial, let's look at this queue
  \item This has amortized O(1) performance cost.
  \item This queue is lazy as long as you don't empty the front, then it is
        fully lazy in the structure of the back list. It only forces the spine
        of the back list when the front list is emptied.
}

@@

\begin{frame}[fragile]
  \frametitle{Recalling \texttt{deQ}}
  \begin{minted}{ocaml}
type 'a lazyList =
  | Cons of ('a Lazy.t *
             'a lazyList Lazy.t)
  | Nil

@@

\note[itemize]{
  \item Having demonstrated how to reify demand behaviors into value, we now
        need to figure out how to write down a specification to check if whether
        a particular run of the program has the specified laziness according to
        the specification
  \item Specifications go backwards from demands on the output to demands on
        the inputs of the function. This is because how much input is demanded
        depends on how much output is demanded, hence the arrows go the other
        way.
  \item Specifications are passed in the actual values of the inputs because
        their demand behaviors can be dependent on those inputs
  \item spineStrictAs takes a lazy data structure, and unions a spine strict
        demand corresponding to the whole structure with another demand on some
        sub-part of that structure
}

>>>>

\newif\ifFinalProject
\FinalProjecttrue

\newif\ifNotFinalProject
\ifFinalProject\NotFinalProjectfalse\else\NotFinalProjecttrue\fi

@@

% Slides for CIS700 Final Presentation
\ifFinalProject
\begin{frame}[fragile]
\frametitle{Lazy Evaluation?}
\begin{itemize}
\item<1->[]
\begin{minted}{haskell}
fibonacci =
  0 : 1 : (zipWith (+) fibonacci (tail fibonacci))
\end{minted}
\item<2->[]
\begin{minted}{haskell}
ghci> fibonacci !! 10
ghci> 55  
\end{minted}
\item<3->[]
\begin{minted}{haskell}
ghci> fibonacci !! 11
ghci> 89
\end{minted}
\end{itemize}
\end{frame}

\note[itemize]{
\item Let's start with an example
\item \texttt{fibonacci} builds a Haskell list, where \text{:} is list cons constructor
\item \texttt{fibonacci} builds a circular structure that contains a reference to itself
\item This is an infinite list
\item but its definition doesn't cause the interpreter to go to an infinite loop
\item Core idea is: code doesn't execute unless output is used
}

\begin{frame}[fragile]
  \frametitle{How to \textit{demand} a lazy value?}
  \centering
  \begin{figure}
    \includegraphics[width=0.75\textwidth]{binary-tree}
  \end{figure}
\end{frame}

\begin{frame}[fragile]
  \centering
  \begin{figure}
    \includegraphics[width=0.75\textwidth]{binary-tree-filled-1}
  \end{figure}
\end{frame}

\begin{frame}[fragile]
  \centering
  \begin{figure}
    \includegraphics[width=0.75\textwidth]{binary-tree-filled-2}
  \end{figure}
\end{frame}

\begin{frame}[fragile]
  \centering
  \begin{figure}
    \includegraphics[width=0.75\textwidth]{binary-tree-filled-3}
  \end{figure}
\end{frame}

\note[itemize]{
\item A lazy value can be thought of as a pointer graph with regions that have
  not matierialized/evaluated yet
\item In this graph, we use white regions to represent parts of the value that
  have not been evaluated yet
\item We evaluate more of the value by following the pointers
}

\begin{frame}[fragile]
\frametitle{Pattern Matching}
\begin{minted}{haskell}
tail :: [Int] -> [Int]
tail xs = case xs of
            [] -> []
            (_ : xs) -> xs
\end{minted}
\end{frame}

\note[itemize]{
\item In a functional language like Haskell, we traverse the pointer graphs of
  lazy values by performing pattern matching on them
\item In the \texttt{tail} function for lists, we evaluate the input list to its
  first constructor
\item Useful: think about Randoop, it iteratively generates new API calls, and
  alternates between generating and testing.
\item If Randoop were able to generate these call sequences lazily, then from
  Randoop developer's perspective there can just be a module that focuses on
  generation and one that focuses on testing
\item Since unused generated API calls are not actually evaluated, it doesn't
  cost additional CPU resource
}

% End of slides for CIS700 Final Presentation
\fi

@@

\ifNotFinalProject
\begin{frame}[fragile]
\frametitle{Lazy lists}
\begin{minted}{ocaml}
type 'a lazyList =
  | Cons of ('a Lazy.t *
             'a lazyList Lazy.t)
  | Nil
\end{minted}
\end{frame}

@@

\note[itemize]{
  \item As a micro example to motivate why thinking about laziness can be
        non-trivial, let's look at this queue
  \item This has amortized O(1) performance cost.
  \item This queue is lazy as long as you don't empty the front, then it is
        fully lazy in the structure of the back list. It only forces the spine
        of the back list when the front list is emptied.
}
\fi

@@

\ifNotFinalProject
\begin{frame}[fragile]
  \frametitle{Recalling \texttt{deQ}}
  \begin{minted}{ocaml}
type 'a lazyList =
  | Cons of ('a Lazy.t *
             'a lazyList Lazy.t)
  | Nil

@@

\note[itemize]{
  \item Having demonstrated how to reify demand behaviors into value, we now
        need to figure out how to write down a specification to check if whether
        a particular run of the program has the specified laziness according to
        the specification
  \item Specifications go backwards from demands on the output to demands on
        the inputs of the function. This is because how much input is demanded
        depends on how much output is demanded, hence the arrows go the other
        way.
  \item Specifications are passed in the actual values of the inputs because
        their demand behaviors can be dependent on those inputs
  \item spineStrictAs takes a lazy data structure, and unions a spine strict
        demand corresponding to the whole structure with another demand on some
        sub-part of that structure
}
\fi
